---
title: "HW3"
author: "Chaoqun Zhou"
date: 2022-10-14
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 2

## Data Wrangling:
```{r, message = FALSE}
p2 = read_csv(file = "./data/accel_data.csv") %>% 
     janitor::clean_names() %>% 
     pivot_longer(activity_1:activity_1440,
                  names_to = "min_of_the_day",
                  values_to = "activity") %>% 
     mutate(min_of_the_day = str_sub(min_of_the_day, 10),
            day = str_to_lower(day),
            weekend = ifelse(day == "saturday"|day == "saturday", "TRUE", "FALSE"),
            weekend = as.logical(weekend),
            )
```

## Describe the resulting dataset. 
This data set is about five weeks of accelerometer data collected on a 63 year-old male with BMI 25. So far, I have import the data into R. And use `janitor::clean_names()` to clean up variables names. Now, this data set has `r ncol(p2)` variables and `r nrow(p2)` observations. It includes variables: `r ls(p2)`. I created a new logical variable `weekend`. It is true when the `day` variable equals to Sunday or Saturday, and FALSE when the `day` variable equal to weekdays. 

## Create a new table showing total activity for each day
```{r, message = FALSE}
  p2 %>% 
  group_by(week, day) %>% 
  summarise(activity_perday = sum(activity)) %>% 
  pivot_wider(
    names_from = day,
    values_from = activity_perday) %>% 
    knitr::kable(digits = 1)
```
As we can see from the table, there are days where this individual have a lot of activity and days this individual has much less activity. Also, there are two days where this individual doesn't have activity (on Saturday of week 4 and 5). For every minute of the day, the individual only have 1 activity per minutes which probably means this individual doesn't carry this device with him.
 




